%
% $XORP: xorp/docs/user_manual/preface.tex,v 1.10 2008/10/02 21:56:44 bms Exp $
%

\chapter*{Preface}

\vspace{-0.1in}
This User Manual describes the process of configuring and operating a
router running XORP.  At the time of writing, XORP is a
work-in-progress, and is evolving relatively quickly.  We hope this
user manual accurately reflects the functionality available in XORP,
but it is likely to date quite quickly.  An up-to-date copy of this
manual will always be available from {\it http://www.candelatech.com/xorp.ct/}.

\section*{Contributing to this Manual}
\vspace{-0.1in}
XORP is an open-source project, and this manual is an open-source
manual.  Like the XORP software, it is covered by the XORP license,
which permits you to modify it and use it for any purpose whatsoever,
so long as the copyright is preserved.  Please help us improve this
manual by sending contributions, suggestions, and criticism to {\it
  feedback@xorp.org}.

\section*{The XORP License}
\vspace{-0.3in}
\begin{alltt}
\small\noindent
\copyright 2004-2011 XORP, Inc and Others
\copyright 2004-2005 University College London

With the exception of code derived from other sources, all XORP software
is copyrighted by XORP, Inc. [Copyright (c) 2001-2011 XORP, Inc and Others].
Files containing derived software are listed in the "LICENSE.other" file
together with their corresponding copyrights and original licenses.

All XORP software is licensed under the GNU General Public License,
Version 2, June 1991 contained in the "LICENSE.gpl" file unless
otherwise indicated in the source file.

Software in source files that refer to the "LICENSE.lgpl" file is
licensed under the GNU Lesser General Public License, Version 2.1,
February 1999 contained in "LICENSE.lgpl".
\end{alltt}

Portions of the XORP CLI use modified version of the {\it libtecla}
library which is covered by the following license.

\section*{The Libtecla License}
\vspace{-0.3in}
\begin{alltt}
\small\noindent
Copyright (c) 2000, 2001 by Martin C. Shepherd.

All rights reserved.

Permission is hereby granted, free of charge, to any person obtaining a
copy of this software and associated documentation files (the
"Software"), to deal in the Software without restriction, including
without limitation the rights to use, copy, modify, merge, publish,
distribute, and/or sell copies of the Software, and to permit persons
to whom the Software is furnished to do so, provided that the above
copyright notice(s) and this permission notice appear in all copies of
the Software and that both the above copyright notice(s) and this
permission notice appear in supporting documentation.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS
OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT
OF THIRD PARTY RIGHTS. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR
HOLDERS INCLUDED IN THIS NOTICE BE LIABLE FOR ANY CLAIM, OR ANY SPECIAL
INDIRECT OR CONSEQUENTIAL DAMAGES, OR ANY DAMAGES WHATSOEVER RESULTING
FROM LOSS OF USE, DATA OR PROFITS, WHETHER IN AN ACTION OF CONTRACT,
NEGLIGENCE OR OTHER TORTIOUS ACTION, ARISING OUT OF OR IN CONNECTION
WITH THE USE OR PERFORMANCE OF THIS SOFTWARE.

Except as contained in this notice, the name of a copyright holder
shall not be used in advertising or otherwise to promote the sale, use
or other dealings in this Software without prior written authorization
of the copyright holder.
\end{alltt}
