%
% $XORP: xorp/docs/olsr/preface.tex,v 1.3 2008/07/23 08:34:36 bms Exp $
%

\chapter*{Preface}

\vspace{-0.1in}
This User Manual describes the process of configuring and operating
the OLSRv1 protocol for XORP.

\section*{Contributing to this Manual}
\vspace{-0.1in}
XORP is an open-source project, and this manual is an open-source
manual.  Like the XORP software, it is covered by the XORP license,
which permits you to modify it and use it for any purpose whatsoever,
so long as the copyright is preserved.  Please help us improve this
manual by sending contributions, suggestions, and criticism to {\it
  feedback@xorp.org}.

\section*{The XORP License}
\vspace{-0.3in}
\begin{alltt}
\small\noindent
\copyright 2004-2008 XORP, Inc.

This program is free software; you can redistribute it and/or modify
it under the terms of the GNU General Public License, Version 2, June
1991 as published by the Free Software Foundation. Redistribution
and/or modification of this program under the terms of any other
version of the GNU General Public License is not permitted.

This program is distributed in the hope that it will be useful, but
WITHOUT ANY WARRANTY; without even the implied warranty of
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. For more details,
see the GNU General Public License, Version 2, a copy of which can be
found in the XORP LICENSE.gpl file.

XORP Inc, 2953 Bunker Hill Lane, Suite 204, Santa Clara, CA 95054, USA;
http://xorp.net
\end{alltt}

Portions of the XORP CLI use modified version of the {\it libtecla}
library which is covered by the following license.

\section*{The Libtecla License}
\vspace{-0.3in}
\begin{alltt}
\small\noindent
Copyright (c) 2000, 2001 by Martin C. Shepherd.

All rights reserved.

Permission is hereby granted, free of charge, to any person obtaining a
copy of this software and associated documentation files (the
"Software"), to deal in the Software without restriction, including
without limitation the rights to use, copy, modify, merge, publish,
distribute, and/or sell copies of the Software, and to permit persons
to whom the Software is furnished to do so, provided that the above
copyright notice(s) and this permission notice appear in all copies of
the Software and that both the above copyright notice(s) and this
permission notice appear in supporting documentation.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS
OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT
OF THIRD PARTY RIGHTS. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR
HOLDERS INCLUDED IN THIS NOTICE BE LIABLE FOR ANY CLAIM, OR ANY SPECIAL
INDIRECT OR CONSEQUENTIAL DAMAGES, OR ANY DAMAGES WHATSOEVER RESULTING
FROM LOSS OF USE, DATA OR PROFITS, WHETHER IN AN ACTION OF CONTRACT,
NEGLIGENCE OR OTHER TORTIOUS ACTION, ARISING OUT OF OR IN CONNECTION
WITH THE USE OR PERFORMANCE OF THIS SOFTWARE.

Except as contained in this notice, the name of a copyright holder
shall not be used in advertising or otherwise to promote the sale, use
or other dealings in this Software without prior written authorization
of the copyright holder.
\end{alltt}
