\chapter*{Preface}

This User Manual describes the process of configuring and operating a
router running XORP Release 1.0.  At the time of writing, XORP is a
work-in-progress, and is evolving relatively quickly.  We hope this
user manual accurately reflects the functionality available in XORP,
but it is likely to date quite quickly.  An up-to-date copy of this
manual will always be available from {\it http://www.xorp.org}.

\section*{Contributing to this Manual}

XORP is an open-source project, and this manual is an open-source
manual.  Like the XORP software, it is covered by the XORP license,
which permits you to modify it and use it for any purpose whatsoever,
so long as the copyright is preserved.  Please help us improve this
manual by sending contributions, suggestions, and criticism to {\it
  feedback@xorp.org}.

\section*{The XORP License}
\vspace{-0.1in}
\begin{alltt}
\small\noindent
\copyright International Computer Science Institute, 2004
\copyright University College London, 2004

Permission is hereby granted, free of charge, to any person obtaining a
copy of this software and associated documentation files (the "Software"),
to deal in the Software without restriction, including without limitation
the rights to use, copy, modify, merge, publish, distribute, sublicense,
and/or sell copies of the Software, and to permit persons to whom the
Software is furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in
all copies or substantial portions of the Software.

The names and trademarks of copyright holders may not be used in
advertising or publicity pertaining to the software without specific
prior permission. Title to copyright in this software and any associated
documentation will at all times remain with the copyright holders.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING
FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER
DEALINGS IN THE SOFTWARE.
\end{alltt}