%
% $XORP: xorp/docs/olsr/preface.tex,v 1.2 2008/07/23 05:10:03 pavlin Exp $
%

\chapter*{Preface}

\vspace{-0.1in}
This User Manual describes the process of configuring and operating
the OLSRv1 protocol for XORP.

\section*{Contributing to this Manual}
\vspace{-0.1in}
XORP is an open-source project, and this manual is an open-source
manual.  Like the XORP software, it is covered by the XORP license,
which permits you to modify it and use it for any purpose whatsoever,
so long as the copyright is preserved.  Please help us improve this
manual by sending contributions, suggestions, and criticism to {\it
  feedback@xorp.org}.

\section*{The XORP License}
\vspace{-0.3in}
\begin{alltt}
\small\noindent
\copyright 2004-2008 XORP, Inc.

Permission is hereby granted, free of charge, to any person obtaining a
copy of this software and associated documentation files (the "Software"),
to deal in the Software without restriction, including without limitation
the rights to use, copy, modify, merge, publish, distribute, sublicense,
and/or sell copies of the Software, and to permit persons to whom the
Software is furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in
all copies or substantial portions of the Software.

The names and trademarks of copyright holders may not be used in
advertising or publicity pertaining to the software without specific
prior permission. Title to copyright in this software and any associated
documentation will at all times remain with the copyright holders.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING
FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER
DEALINGS IN THE SOFTWARE.
\end{alltt}

Portions of the XORP CLI use modified version of the {\it libtecla}
library which is covered by the following license.

\section*{The Libtecla License}
\vspace{-0.3in}
\begin{alltt}
\small\noindent
Copyright (c) 2000, 2001 by Martin C. Shepherd.

All rights reserved.

Permission is hereby granted, free of charge, to any person obtaining a
copy of this software and associated documentation files (the
"Software"), to deal in the Software without restriction, including
without limitation the rights to use, copy, modify, merge, publish,
distribute, and/or sell copies of the Software, and to permit persons
to whom the Software is furnished to do so, provided that the above
copyright notice(s) and this permission notice appear in all copies of
the Software and that both the above copyright notice(s) and this
permission notice appear in supporting documentation.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS
OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT
OF THIRD PARTY RIGHTS. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR
HOLDERS INCLUDED IN THIS NOTICE BE LIABLE FOR ANY CLAIM, OR ANY SPECIAL
INDIRECT OR CONSEQUENTIAL DAMAGES, OR ANY DAMAGES WHATSOEVER RESULTING
FROM LOSS OF USE, DATA OR PROFITS, WHETHER IN AN ACTION OF CONTRACT,
NEGLIGENCE OR OTHER TORTIOUS ACTION, ARISING OUT OF OR IN CONNECTION
WITH THE USE OR PERFORMANCE OF THIS SOFTWARE.

Except as contained in this notice, the name of a copyright holder
shall not be used in advertising or otherwise to promote the sale, use
or other dealings in this Software without prior written authorization
of the copyright holder.
\end{alltt}
